Equilivest: A Robotic Vest to aid in Post-Stroke Dynamic Balance Rehabilitation

### 2022-09-16 iros.tex
Almost the finished version.

\title{Equilivest: A Robotic Vest to aid in Post-Stroke Human Balance Rehabilitation
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Franco Paviotti}
\IEEEauthorblockA{\textit{Bioingeniería} \\
\textit{Instituto Tecnológico de Buenos Aires}\\
Buenos Aires, Argentina \\
fpaviotti@itba.edu.ar}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Esteban Buniak}
\IEEEauthorblockA{\textit{Ingeniería en Informática} \\
\textit{Instituto Tecnológico de Buenos Aires}\\
Buenos Aires, Argentina \\
ebuniak@itba.edu.ar}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Rodrigo Ramele}
\IEEEauthorblockA{\textit{Ingeniería en Informática} \\
\textit{Instituto Tecnológico de Buenos Aires}\\
Buenos Aires, Argentina \\
rramele@itba.edu.ar}
\and
\IEEEauthorblockN{4\textsuperscript{th} Juan Miguel Santos}
\IEEEauthorblockA{\textit{Ingeniería en Informática} \\
\textit{UNAHUR}\\
Hurlingham, Argentina \\
jsantos@gmail.com}
\and
\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{Ingeniería en Informática} \\
\textit{Instituto Tecnológico de Buenos Aires}\\
Buenos Aires, Argentina \\
fpaviotti@itba.edu.ar}
\and
\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{Ingeniería en Informática} \\
\textit{Instituto Tecnológico de Buenos Aires}\\
Buenos Aires, Argentina \\
fpaviotti@itba.edu.ar}
}

@@

\begin{abstract}
Stroke is a devastanting disease that affects bla bla bla.
It has been shown that neurofeedback can help patients to regain lost motor activity and recover functionality.
In this work, we are proposing a device that help patients by augmentation sensory sensation and pinning information to certain patterns in gait.  We are presenting results from surveys to patient's caregivers, patients and medical personal.  We are presenting the experimentation paradigm, and the robotic device that we created to help the patient.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

@@

Brain stroke is a devastating disease.  It affects worldwide population and is the main cause of disabilities worldwide.  Disabilities related to stroke can affect motor pathways, and may lead to several motor disability disorders.  Hence, as early as possible the rehabilitation can be put into place, better the outcome that can be obtained by the treatment.

Strong evidence suggests that neuroplasticity can be enhanced by neural rehabilitation~\cite{Albert.etal2012}.  These procedures are aimed to relearn movements that can trigger new pathway generation which reroute, or even completely replace, those pathways that were damaged by the stroke.  Neurorehabilitation procedures are performed by a group of interdisciplinary care givers and technicians.  Recently, biofeedback techniques, aiming at providing extra information to the patient that can be used to aid in the relearning has appear as an alternative treatment to increase neuralplasticitly, in the form of Wearable devices-based biofeedback rehabilitation (WDBR) home rehabilitation, or even daily use\cite{Peake.etal2018}.  Authors found evidence that biofeedback can indeed help rehabilitation procedures.

This work relies on the following idea:  the addition of an independent and new signal, that can be assimilated as an extrasensory input, could  provides extra information which could improve dynamic balance performance on stroke patients which may have yet insufficiency to deal properly with walking.  The idea is that we have now the ability to provide meaningful balance information with hardware and software that can provide this extra signal in the form of any form of stimulation, for instance vibrotactile.

We are presenting here the development of a device which is grounded on this idea, and is aiming to help a post-stroke patient with a dynamic balance problem as a case study.  This is implemented with a smart-vest~\cite{Brandebusemeyer.etal2021}, which we will call, equilivest, that can detect some particular situation and implement in a way to fix one particular problem.  We aim to provide motor learning, which means that you need to provide a fading vibrotactile feedback, and that the signal needs to be provided in a way which is less conscious as possible.  The device is based on the idea to promote plasticity by giving her vibrotactile stimulation based on kinematic and dynamics measurements.  The idea is to exploits these ideas from neurofeedback that we can create "external" sensors by providing some form of vibrotactile stimulation on her belly and link that stimulation to some form of measurements from IMUs (inertial sensors).

The form of validation procedures that is required in clinical settings\cite{Papastylianou.etal2016}.

Section presents the case study.  Next section presents the underlying clinical hypothesis that this work is trying to promote.  Finally section this and that present the experimental design.  Results are presented on the next section, and this work concludes with discussion and conclussion.

@@

\subsection{Single Patient Case Study}

Patient is a 31 years old female, perfectly healthy, who suffered an acute brainstem stroke after giving birth  The stroke was on posterior fossa subarachnoid due to a brain artereovenous malformation (AVM), which was somehow facilitated by pregnancy or puerperium~\cite{Porras.etal2017}.  Patient was in coma for around 2 months, after that unable to walk, move, talk or swallow.  After two years of intensive rehabilitation, Patient managed to recover significantly, including from dysphagia, which were very important in order to remove the feeding tube allowing her to start speech recovering.

After 24 months since event, the patient, was discharged from hospital and only maintained a 3 times per week rehabilitation treatment, focusing on a remaining affection related with dynamic walking balance problems.  Patient achieved higher index scores in static balance tests and is fully able to hip-balance and ankle-balance.  Patient has recovered muscle in her legs and can perform lower-limb workouts without problems.  Her vision is normal.  Somatosensorial system is optimal on foot sole.  Tests performed to verify proprioceptive system were successfully passed.

However, when Patient tries to walk on open-spaces, or with confronting lights (like walking towards sunlight) or with other people moving around, she is unable to keep up with the pace of the gait and falls frequently.  This is consistent, under these three situations.  Nowadays, the patient can walk with a Canadian cane or a non-actuated walking helper which she is hesitant to use, both.

\subsection{Underlying hypothesis}

Human balance is composed of a complex interaction of different subsystems, which includes somatosensorial information, vestibular system, visual information, which is later processed in different networks of the Central Nervous System, and actuated by motor pathways at many different scales.

Tables ~\ref{tab:patientsurvey} and \ref{tab:caregiverssurvey} show the results of the surveys performed on the Patient and their main Rehabilitaion caregivers.  Using the results from the surveys, we postulate three different systems that could benefit from the introduction of an external signal in the form of biofeedback that can aid in rehabilitation procedures.

@@

\begin{itemize}
\item Vestibular information on fusion of vestibular information
\item Bradykinesia: the processing speed required to effectively perform the processing and actuation is not enough.
\item Unusual gait
\end{itemize}

Based on the surveys and the information provided by caregivers there seems not to be any somatosensorial deficit\cite{cite}.

\subsection{Robotic Device Vest}
Description of the system.
How does it work, how is the detection of IMU working
Internet of Robotic Things.
Arduino

\textbf{Why only vibrotactile stimulation and not auditory stimulation}

\subsection{Experimental Design}

Three experiments were designed in order to test each one of the hypothesis.  All the experiments were approved by the ITBA University ethical committee.  Participants were recruited voluntarily.

@@

\subsubsection{Vestibular Feedback}

The underlying idea is that signals from IMU can be feed back to the patient by vibrotactile stimulation on their belly.  The hypothesis is that if there is a vestibular problem that forbids the patient to receive or evaluate the vestibular information appropriately, this external signal can be available for the patient to integrate it into the dynamic balance integration.

In order to test it, 5 healthy participants are going to be recruited to simulate falling breakpoint conditions on the IMU information.  Participants wear the vest and perform a forward walking exercise with the upper-trunk leaned forward at different angles progresively until they can no longer cope with the unbalance situation without stepping forward.

Afterward, the experiment will be repeated with the 5 participants activating the vibrotactile stimulation which will map progressively the inclination angle.  Hence it will provide an extra signal that can give a accurate information in relation to the stability of the upper-trunk in relation to the walking gait.

\subsubsection{Bradykinesia: Fall prediction due to insufficient speed of step}

All the IMU information represent a multichannel time series.  Hence, it can be used to predict a falling situation.

In this case, 5 healthy participants perform 20 controlled falls over gym mats.

(3) Ver en base a la encuesta como se da la situación de caída más común que Paula relata, en base a lo que ella manifiesta.  Entonces lo que hacemos es tomamos 5 sujetos y los hacemos caer 5 veces cada uno sobre la colchoneta (o 10 o 20 x sujeto, lo que más puedas).  Con estos datos, que son series de tiempo de 6 variables, intetamos ver si con un modelo de ML podemos predecir el momento de la caída simulada de cada uno de ellos y ANTICIPARNOS .  La idea entonces es que suene el motor cuando el sistema detecta con antelación que puede venir una caída inminente.  Aca se muestra el accuracy en la detección de la caída en los datos recabados.

10 participants simulating falls.

\subsubsection{Gait Pacemaker}

This is a gait pacemaker coupled with a podometer.  It has been showed that gait synchronization with music achieved better effect \cite{Roerdink.etal2007}.

(4) En este caso, lo que hacemos es armar un gráfico de gait y lo que hace el dispositivo es detectar un punto en el ciclo y vibrar de forma recurrente, como un marca paso, basado en un podómetro.  En este caso, tomás 5 personas las hacés caminar, determinás cómo son esas curvas y accionás el motor siempre en el mismo momento.  Lo que mostramos son los gráficos y el momento donde se dispara la estimulación.

%\subsection{Abbreviations and Acronyms}\label{AA}
%Define abbreviations and acronyms the first time they are used in the text, 
%even after they have been defined in the abstract. Abbreviations such as 
%IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
%abbreviations in the title or heads unless they are unavoidable.
%
%\subsection{Units}
%\begin{itemize}
%\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
%\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
%\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
%\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
%\end{itemize}
%
%\subsection{Equations}
%Number equations consecutively. To make your 
%equations more compact, you may use the solidus (~/~), the exp function, or 
%appropriate exponents. Italicize Roman symbols for quantities and variables, 
%but not Greek symbols. Use a long dash rather than a hyphen for a minus 
%sign. Punctuate equations with commas or periods when they are part of a 
%sentence, as in:
%\begin{equation}
%a+b=\gamma\label{eq}
%\end{equation}
%
%Be sure that the 
%symbols in your equation have been defined before or immediately following 
%the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
%the beginning of a sentence: ``Equation \eqref{eq} is . . .''
%
%\subsection{\LaTeX-Specific Advice}
%
%Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
%of ``hard'' references (e.g., \verb|(1)|). That will make it possible
%to combine sections, add equations, or change the order of figures or
%citations without having to go through the file line by line.
%
%Please don't use the \verb|{eqnarray}| equation environment. Use
%\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
%environment leaves unsightly spaces around relation symbols.
%
%Please note that the \verb|{subequations}| environment in {\LaTeX}
%will increment the main equation counter even when there are no
%equation numbers displayed. If you forget that, you might write an
%article in which the equation numbers skip from (17) to (20), causing
%the copy editors to wonder if you've discovered a new method of
%counting.
%
%{\BibTeX} does not work by magic. It doesn't get the bibliographic
%data from thin air but from .bib files. If you use {\BibTeX} to produce a
%bibliography you must send the .bib files. 
%
%{\LaTeX} can't read your mind. If you assign the same label to a
%subsubsection and a table, you might find that Table I has been cross
%referenced as Table IV-B3. 
%
%{\LaTeX} does not have precognitive abilities. If you put a
%\verb|\label| command before the command that updates the counter it's
%supposed to be using, the label will pick up the last counter to be
%cross referenced instead. In particular, a \verb|\label| command
%should not go before the caption of a figure or a table.
%
%Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
%will not stop equation numbers inside \verb|{array}| (there won't be
%any anyway) and it might stop a wanted equation number in the
%surrounding equation.
%
%\subsection{Some Common Mistakes}\label{SCM}
%\begin{itemize}
%\item The word ``data'' is plural, not singular.
%\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
%\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)

>>>>

\title{Equilivest: A Robotic Vest to aid in Post-Stroke Dynamic Balance Rehabilitation
}

\author{%
Franco Paviotti$^{1}$, Esteban Buniak$^{2}$, Rodrigo Ramele$^{3}$, Orestes Freixes$^{4}$ and Juan Miguel Santos$^{5}$% <-this % stops a space
\thanks{$^{1}$F. Paviotti is with Bioengineering, Instituto Tecnológico de Buenos Aires
                    Buenos Aires, Argentina
        {\tt\small fpaviotti@itba.edu.ar}}%
\thanks{$^{2}$E.Buniak is with Ingeniería en Informática, Instituto Tecnológico de Buenos Aires
                    Buenos Aires, Argentina
        {\tt\small ebuniak@itba.edu.ar}}%
\thanks{$^{3}$R.Ramele is with Ingeniería en Informática, Instituto Tecnológico de Buenos Aires
                    Buenos Aires, Argentina
        {\tt\small rramele@itba.edu.ar}}%
\thanks{$^{4}$O. Freixes is with the CINER Centro Integral de Neurorehabilitación,
                    Buenos Aires, Argentina
        {\tt\small orestesfreixes@gmail.com}}%
\thanks{$^{5}$J.M. Santos is with the UNAHUR Universidad Nacional de Hurlingham,
                    Buenos Aires, Argentina
        {\tt\small juanmiguelsantos@gmail.com}}%
%\thanks{*This work was not supported by any organization}% <-this % stops a space}
}

%\author{\IEEEauthorblockN{1\textsuperscript{st} Franco Paviotti}
%\IEEEauthorblockA{\textit{Bioingeniería} \\
%\textit{Instituto Tecnológico de Buenos Aires}\\
%Buenos Aires, Argentina \\
%fpaviotti@itba.edu.ar}
%\and
%\IEEEauthorblockN{2\textsuperscript{nd} Esteban Buniak}
%\IEEEauthorblockA{\textit{Ingeniería en Informática} \\
%\textit{Instituto Tecnológico de Buenos Aires}\\
%Buenos Aires, Argentina \\
%ebuniak@itba.edu.ar}
%\and
%\IEEEauthorblockN{3\textsuperscript{rd} Rodrigo Ramele}
%\IEEEauthorblockA{\textit{Ingeniería en Informática} \\
%\textit{Instituto Tecnológico de Buenos Aires}\\
%Buenos Aires, Argentina \\
%rramele@itba.edu.ar}
%\and
%\IEEEauthorblockN{4\textsuperscript{th} Juan Miguel Santos}
%\IEEEauthorblockN{\textit{Ingeniería en Informática} \\
%\textit{UNAHUR}\\
%Hurlingham, Argentina \\
%jsantos@gmail.com}
%\and
%\IEEEauthorblockN{5\textsuperscript{th} Orestes Freixes}
%\IEEEauthorblockN{\textit{CINER} \\
%\textit{Centro Integral de NeuroRehabilitación}\\
%Buenos Aires, Argentina \\
%orestesfreixes@gmail.com}
%}

@@

\begin{abstract}
Stroke is a medical condition that can affect motor function, particularly dynamic balance.  Biofeedback can aid in rehabilitaiton procedures which help patients to regain lost motor activity and recover functionality.  In this work, we are presenting a smart-vest device that assist in rehabilitation procedures by providing feedback in the form of vibrotactile stimulation. Information provided by principal caregivers, family, patient in the form of surveys and interview, is used to derive potential clinical causal hypothesis and these are used to drive the experimentation paradigm, and the robotic smart-vest to aid in the whole procedure...
\end{abstract}

\begin{IEEEkeywords}
Stroke, Balance, Rehabilitation, Biofeedback, Vibrotactile
\end{IEEEkeywords}

@@

Brain stroke is a devastating medical condition, that affects world population and is the main cause of disabilities worldwide~\cite{Caplan.etal2023}.  Disabilities related to stroke can affect motor pathways, and may lead to several motor function disorders.   One important aspect of motor function is balance which is the ability to control the center of mass of the body inside the base support provided by the lower limb~\cite{Bowman2021}.  Stroke can affect dynamic balance as well,  which is manifested while walking, impairing autonomy and independence, important factors in Activities of Daily Living (ADL) particularly for young patients~\cite{Afrasiabifar.etal2020,Donato.etal2016}.

% As early as possible the rehabilitation can be put into place, better the outcome that can be obtained by the treatment.

Strong evidence suggests that neuroplasticity can be enhanced by neural rehabilitation~\cite{DeAngelis.etal2021,Albert.etal2012}.  These procedures are aimed to relearn movements that can trigger new neural pathway generation which reroute, or even completely replace, those pathways that were damaged by the stroke.  Neurorehabilitation procedures are performed by a group of interdisciplinary caregivers and technicians.  Recently, biofeedback techniques, aiming at providing extra information to the patient that can be used to aid in the relearning have appeared as an alternative treatment to increase neuralplasticity. These are in the form of Wearable devices-based biofeedback rehabilitation (WDBR)~\cite{Peake.etal2018} or more advanced robotic rehabilitation gait devices~\cite{Zhao.etal2022,Peshkin.etal2005,Tong.etal2006}.

%This published evidence emphasizes that this biofeedback can indeed enhance the results of rehabilitation procedures.

Therefore, the addition of an independent and new peripheral therapeutic signal, that can be assimilated as an extrasensory input~\cite{Brandebusemeyer.etal2021},  could improve dynamic balance performance on post-stroke patients which may have yet insufficiency to deal properly with the complexities of walking.  Meaningful balance information, in terms of timing and location, can provide this extra signal in any form of stimulation, particularly vibrotactile feedback (VF).  Although the effectiveness of biofeedback on static balance has been studied more extensively in the literature~\cite{DeAngelis.etal2021}, to the best of our knowledge works dealing with dynamic balance problems while walking have been more scarce.

This work presents the development of a device which is grounded on this idea, and aims to help a post-stroke patient with a remaining dynamic balance problem, presenting it as a case study.  The proposed development is implemented as a smart-vest~\cite{Brandebusemeyer.etal2021}, which we will call, \textit{Equilivest}, that address three possible clinical hypothesis of the underlying problem.  We aim to provide motor learning, meaning to generate a fading compliant assist-as-needed vibrotactile feedback signal which is manifested to the patient as less conscious as possible~\cite{Srivastava.etal2016,Donato.etal2016}.  The device aims to promote plasticity by producing timing vibrotactile stimulation based on kinematic and dynamics measurements.

%The idea is to exploits these ideas from neurofeedback that we can create "external" sensors by providing some form of vibrotactile stimulation on her belly and link that stimulation to some form of measurements from IMUs (inertial sensors).

%The form of validation procedures that is required in clinical settings\cite{Papastylianou.etal2016}.

Section \ref{sec:case-study} presents the case study.  Next section \ref{sec:hypothesis} summarize the results of the interviews and surveys performed by the patient, their family and professional caregivers, presenting the underlying clinical causal hypothesis.  Section \ref{sec:vest} describes the vest details and architecture.  The experimental design protocol is expounded in Section \ref{sec:experimental}.  Preliminary results are presented on the next section, and this work concludes with discussion and conclusion.

@@

\subsection{Patient Case Study}
\label{sec:case-study}

Patient is a 31 years old female, without any history of chronic ailment, who suffered an acute brainstem stroke after giving birth.  The stroke was on posterior fossa subarachnoid due to a brain artereovenous malformation (AVM), which could be linked to pregnancy or puerperium~\cite{Porras.etal2017}.  Patient was in coma for around 2 months, and after that unable to walk, move, talk or swallow.  After two years of intensive rehabilitation, Patient managed to recover significantly, including from dysphagia, which was very important in order to remove the feeding tube allowing her to start speech recovering.

After 24 months since event, the patient, was discharged from hospital and only maintained a 3 times per week rehabilitation treatment, focusing on a remaining affection related with dynamic balance problems while walking.  The patient achieved satisfactory index scores in static balance tests and is able to perform static hip-balance and ankle-balance.  She has recovered muscle in her legs and can perform lower-limb exercises.  Her vision is normal.

However, when the patient tries to walk on open-spaces, or with confronting lights (like walking towards sunlight), with other people moving around, or while walking concentration fades, she is unable to keep up with the pace of the gait and the risk of fall increases.  Nowadays, the patient walks freely unaided at home but requires a Canadian cane otherwise.

%\begin{table*}[t]
%\begin{center}
%\begin{tabular}[!t]{|c|c|}
%\hline
%Question & Avg. answer \\
%\hline
%Podes describir cómo y cuándo se expresa la perdida del equilibrio y su proceso? - Esta pregunta nos servirá para determinar con mayor precisión los métodos de sensado y alarma a desarrollar  & Cuando estoy cansada, distraída o muy nerviosa\\
%\hline
%¿Siempre se desarrolla de la misma manera? - Con esta pregunta queremos profundizar sobre la anterior con el objetivo de una mejor comprensión &  Si \\
%\hline
%¿Cuál? Si la respuesta fue no, deje sin contestar  & Me desequilibro\\
%\hline
%¿Podés evidenciar cuando estas por empezar a perder el equilibrio? - Con esta pregunta intentamos conocer un fenómeno que permita detectar el principio de caída & Si \\
%\hline
%Si la respuesta fue si ¿Qué sentís? & Un poco no siempre siento inestabilidad previo  \\
%\hline
%¿Con qué frecuencia usas el andador? - Puede seleccionar mas de una opción. Así nos ayudará a entender mejor el cuadro & Solo fuera de casa \\
%\hline
%¿Existe algún denominador común en las caídas? - La presencia de este sería importantísima para  un fenómeno o señal a medir y caracterizar que sería muy útil para la detección del principio de caídas. Si hay algo que siempre se repite podríamos enfocarnos en detectar eso. & Si \\
%\hline\\
%¿Cuál? Si la respuesta fue "No" deje la respuesta en blanco & Lo que anteriormente dije estoy cansada o distraída o nerviosa  \\
%\hline
%¿Cómo estás trabajando la recuperación hoy por hoy? - Con esta pregunta buscamos conocer el estado de situación y trabajo actual & Fortalecimiento de músculos   \\
%\hline
%¿Hay algo más que desees comentar o expresar? & No uso andador es un bastón canadiense. 
%Ahora me doy cuenta antes que me voy a caer, pero queda poco para evitarlo.   \\
%\hline
%\end{tabular}
%\caption{table}{Answers to survey questions.}
%\label{tab:alpi_q_table_1}
%\end{center}
%\end{table*}

%\begin{table*}[t]
%\begin{tabular}[!t]{|c|c|}
%Question & Avg. answer \\
%¿Cuál es la razón de la pérdida del equilibrio en la paciente? - Esta pregunta nos permite conocer la opinión de los profesionales e interiorizarnos en el caso.  & Lesión en el tronco del encéfalo\\
%¿Podes describir cómo y cuándo se expresa la perdida del equilibrio y su proceso? - Con la descripción podemos comprender como y en que circunstancias se desencadena el fenómeno de la caída y conocer mejor el caso. &  La pérdida del equilibrio se expresa en el inicio y durante la marcha. cuando se detiene también. Camina con compensación visual y cuando quita la vista el suelo (ej mira para el costado aumenta la inestabilidad) Realiza una marcha sin el automatismo habitual. Le demanda mucha concentracion.  \\
%¿ La paciente puede caminar sin un dispositivo de tipo ayuda marcha? De ser así, ¿Bajo que circunstancias? - Conociendo mejor la situación actual podemos tomar mejores decisiones en cuanto a como posicionar sensores y medir.  & Solo de manera  terapéutica. habitualmente la trabajamos con máxima supervisión.\\
%¿Presenta una pérdida de fuerza muscular que dificulte el caminar? - Esta pregunta permite conocer el caso en mayor detalle  con propiedad para evitar adelantarnos a hipótesis de trabajo erróneas. & No \\
%¿Presenta dismetría? - Nuevamente queremos adelantarnos a hipótesis erróneas a la hora de decidir el método de sensado y procesamiento de las señales. La longitud o duración de los movimientos puede usarse como un indicador para nosotros, como también asi los ritmos. & Si \\
%¿Presenta temblores?  - En caso de presentar temblores estos se verán reflejados en las señales y es útil tenerlo en cuenta ya que caracterizará a nuestra señal. & No \\
%¿Presenta alteraciones en la marcha en instantes previos a la caída? - En caso de existir algún impulso inicial que podamos sensar o medir de forma indirecta podemos anticiparnos a la perdida del equilibrio en las señales. & Si \\
%\hline\\
%¿Cuáles? Si la respuesta fue "No" deje en blanco & Fallas las reacciones de tobillo y cadera \\
%¿Sufrió alteraciones en la sensibilidad profunda? - Existe un trabajo previo en la Universidad de Buenos Aires que usa un método de sensado orientado a la presión de la pisada. Esto puede orientarnos en otro sentido en el sensado y estimulación. & No  \\
%Antes de perder el equilibrio, ¿Trastabilla?¿Acelera la marcha? -  Al igual que en otras preguntas buscamos obtener información extra para el  procesamiento de las señales y detección del principio de caída. Estos fenómenos pueden medirse o detectarse y evidenciarían el inicio de una posible caída. & Otro síntoma  \\
%Si la respuesta fue 'Otro síntoma' ¿Cuál? - En caso contrario deje sin responder & Reacción lenta \\
%¿Presenta algún tipo de marcha patológica? -la caracterización del caso permite decidir formas y métodos de sensado y procesamiento. & Si  \\
%¿Cuál? Si la respuesta fue "No" deje la respuesta en blanco & Marcha Atáxica \\
%¿Existe algún denominador común en las caídas? - La presencia de este sería importantisima para  un fenómeno o señal a medir y caracterizar que sería muy útil para la detección del principio de caídas. Si hay algo que siempre se repite podríamos enfocarnos en detectar eso. & Si \\
%¿Cuál? Si la respuesta fue "No" deje la respuesta en blanco & El denominador es cuando deja de realizar la marcha de manera consciente. Por ejemplo, cuando camina y tiene que realizar alguna tarea que implique procesos cognitivo. Dual Task  \\
%¿Se hizo un estudio de evaluación de marcha? - Esta pregunta podría aportar información suplementaria a preguntas anteriores que nos permiten conocer e interiorizarnos en el caso & No  \\
%¿Se hizo evaluación de riesgo de caídas? - Esta pregunta puede aportar algo de información extra para el conocimiento del caso. & Si  \\
%¿Hay alguna manera en que trabaje en la reeducación de la paciente? -  Quisiéramos conocer un poco mas la situación actual y las actividades que realiza la paciente para su recuperación. & Si a través de la reeducación del VOR,  facilitación del equilibrio estático y dinámico, entrenamiento del automatismo de la marcha con diferentes niveles de dificultad.  \\
%¿Hay algo mas que desee comentar o expresar?  & Mes a mes se evidencia una mejora, todavía no alcanza para que pueda caminar sin supervisión.  \\
%\caption{table}{Answers to survey questions.}
%\label{tab:alpi_q_table_1}
%\end{table*}

\subsection{Underlying hypothesis}
\label{sec:hypothesis}

Human balance is composed of a complex interaction of different subsystems, which includes somatosensorial information, vestibular system and visual information as input sources.  These are later processed in different networks of the Central Nervous System (CNS), and finally actuated by motor pathways at many different scales~\cite{Donato.etal2016}.

\begin{table*}[t]
\begin{center}
\begin{tabular}[!t]{|p{0.45\linewidth} |p{0.45\linewidth} |}
\hline
\textbf{Question} & \textbf{Answer} \\
\hline
\hline
Under which situations do you feel you are prone to fall ? & When I am tired, distracted or stressed. \\
\hline
What do you feel ?  & I feel that I loss my balance\\
\hline
What do you feel before you fall ? & Sometimes I do feel something before.  I realize that I am going to fall, but there isn't anything I can do to avoid it. \\
\hline
When do you use the cane ?  & Only when I am out of home.\\
\hline
\end{tabular}
\vspace{10pt}
\caption{Survey and responses provided by the patient.}
\label{tab:patientsurvey}
\end{center}
\end{table*}

\begin{table*}[t]
\begin{center}
\begin{tabular}[!t]{|p{0.45\linewidth} |p{0.45\linewidth} |}
\hline
\textbf{Question} & \textbf{Answer} \\
\hline
\hline
Describe the situation when the patient loss her balance & Loss of balance is manifested at the beginning, during the gait cycle or with sudden stops. Walking is performed using intense visual compensation.   When the patient is distracted, or when she stops looking at the floor, or looks sideways, instability increases.  Gait  is not automated, and demands cognitive resources. \\
\hline
Is there any significant muscle loss in lower-limbs ? & No \\
\hline
Does she presents dysmetria ? & No \\
\hline
Does she presents heavy shaking ? & No \\
\hline
Does she have any visible reaction before falling?  & She presents dynamic alterations in hip and ankle compensation.\\
\hline
Does she have any somatosensorial alteration?  &  No \\
\hline
Does she have any proprioceptive alteration?  &  No \\
\hline
Is the patient gait normal ?  &  No, the patient presents ataxic gait, likely triggered by slow reaction to perform lower-limb balance correction. \\
\hline
Is there any behavioural pattern linked to falling events ? &  Yes, cognitive workload.  Dual tasks situations, when the patient needs to perform something extra while walking. \\
\hline
Describe current treatment. & The patient is currently working on rehabilitation exercises to retrain her Vestibulo-Ocular Reflex (VOR), to improve her static and dynamic balance, and improve her locomotor automatism.  She is improving on a montly basis, but still she has not reached the level to walk autonomy without any aid. \\
\hline
\end{tabular}
\vspace{10pt}
\caption{Survey and responses provided by his main professional rehabilitation caregiver.}
\label{tab:caregiverssurvey}
\end{center}
\end{table*}

We perform a series of surveys and interviews with the patient and their caregivers.  Main results are summarized in 
Tables ~\ref{tab:patientsurvey} and \ref{tab:caregiverssurvey}.  Based on the clinical history, the results from the surveys and interviews, we postulate three different potential clinical situations that could use the external signal in the form of biofeedback and could potentially aid in rehabilitation procedures.  The first (i) is a potential problem in the integration of vestibular information while walking, the second (ii) is bradykinesia where the required processing speed to effectively adjust the lower limb to keep the center of mass inside the base support is not achieved.  Finally, the (iii) hypothesis is an ataxic gait, where non-automated gait produces an increase in the likelihood of failing.

@@

%\begin{itemize}
%\item Vestibular information or fusion of vestibular information
%\item Bradykinesia: the processing speed required to effectively perform the processing and actuation is not enough.
%\item Ataxic gait: the movement is not normally regulated.
%\end{itemize}
%

\subsection{Robotic Device Vest}
\label{sec:vest}

The vest system prototype is based on Internet of Robot Thing (IoRT) technology~\cite{Simoens.etal2018,Domingo.etal2012}. The controller is an Arduino UNO (Arduino LLC, Italy) board coupled with ESP8266 shield (WeMos, United Kingdom).  The system is powered by a commercial power bank with Li-Po 18650 cells (Ipower, United States), an inertial measurement unit Inertial Measurement Unit (IMU) MPU 6050 (OEM ITG/MPU6050) and a FA-12350 DC motor scavenged from old compact discs which provide the vibrotactile feedback.  The IMU uses the accelerometer and gyroscope data to get 3-dimensional angular acceleration and measure pitch roll and yaw angles. These angles are calculated from a set of transforming equations and processed with a complementary filter which relies 80\% in gyroscope data and 20\% in accelerometer data~\cite{Fetick.2022}. These values are transmitted by telemetry in real-time as UDP packets for off-board register, processing and further analysis. Data recovered by the device is sampled at 1000 Hz.

The smart-vest prototype can be seen in Figure~\ref{fig:smart-vest} ......

%\textbf{Why only vibrotactile stimulation and not auditory stimulation}

\subsection{Experimental Design}
\label{sec:experimental}

Three experiments are designed in order to assess each one of the potential clinical causes, and to derive for each one of them a stimulation strategy.  Participants are recruited voluntarily and the experiment is conducted anonymously in accordance with the declaration of Helsinki published by the World Health Organization. No monetary compensation is handed out and all participants agree and sign a written informed consent. This study is approved by the Departamento de Investigación y Doctorado, Instituto Tecnológico de Buenos Aires (ITBA).  All the participants wear the smart-vest with an elastic band pressed tightly towards the chest, tied in no-restrictive manner, holding the sensor firmly.

@@

\subsubsection{Vestibular Information Integration}

This experiment aims to study the falling process. Participants are told to lean forward until they couldn't maintain balance anymore and let themselves fall into a mattress. This experiment is used to determine a threshold for the data that can help to identify the breakpoint conditions based on the IMU information where each subject falls.

In order to test it, 5 healthy participants are recruited to perform 10 runs of falling situations  Participants wearing the vest  perform a one-step forward walking exercise with the upper-trunk leaned forward at different angles progressively until they can no longer cope with the unbalance situation and fall to the mattress.

\subsubsection{Ataxic Gait}

This experiment has the purpose of analyzing and studying gait's behavior. In order to accomplish this,  5 participants are instructed to perform 10 sessions of walking in a straight line across ten meters, performing Ten-Meter Walking (10MWT) assesment~\cite{Olmos.etal2008}.  We collected and analyzed pitch, roll and yaw values as well as angular acceleration changes registered by the IMU gyroscope. In this study our goal is to be able to process and identify gait abnormalities as well as all steps occurring and any fall that might develop.

\subsubsection{Bradykinesia}

This experiment involves the coupling of the other two.  Five participants are instructed to perform 10 runs of a ten-meter walking procedure, followed by a falling into a mattress.  The purpose of this experiment is obtain a multichannel time series of the whole sequence, including the walk and the falling action.

%(3) Ver en base a la encuesta como se da la situación de caída más común que Paula relata, en base a lo que ella manifiesta.  Entonces lo que hacemos es tomamos 5 sujetos y los hacemos caer 5 veces cada uno sobre la colchoneta (o 10 o 20 x sujeto, lo que más puedas).  Con estos datos, que son series de tiempo de 6 variables, intetamos ver si con un modelo de ML podemos predecir el momento de la caída simulada de cada uno de ellos y ANTICIPARNOS .  La idea entonces es que suene el motor cuando el sistema detecta con antelación que puede venir una caída inminente.  Aca se muestra el accuracy en la detección de la caída en los datos recabados.

%(4) En este caso, lo que hacemos es armar un gráfico de gait y lo que hace el dispositivo es detectar un punto en el ciclo y vibrar de forma recurrente, como un marca paso, basado en un podómetro.  En este caso, tomás 5 personas las hacés caminar, determinás cómo son esas curvas y accionás el motor siempre en el mismo momento.  Lo que mostramos son los gráficos y el momento donde se dispara la estimulación.
